Asteroseismology of GD358 with complex C/O core profiles

### 2016-08-15 tmp1/msnew.tex
uploading

\begin{abstract}
We report on the analysis of 34 years of photometric observations of the pulsating
helium atmosphere white dwarf GD358.  The complete data set includes archival data from 1982-2006, 
and 1195.2 hours of new observations from 2007-2016. From this data set, we extract 15 frequencies 
representing g-mode pulsation modes, adding 4 modes to the 11 modes 
known previously. We present evidence that these 15 modes are $\ell=1$ modes, 13 of which belong to 
a consecutive sequence in radial overtone
$k$. We perform a detailed asteroseismic analysis using models that include parameterized, complex carbon and oxygen 
core composition profiles to fit the periods. Recent spectroscopic analyses place GD358 near the red edge of the DBV instability strip, at 24,000 $\pm$ 500~K and a \logg of 7.8 $\pm$ 0.08 dex. The surface gravity translates to a mass range of 0.455 to 0.540 \msun. Our best fit model has a temperature of 23,650~K and a mass of 0.5706 \msun ~. That is slightly more massive than what the most recent spectroscopy suggests. We find a pure helium layer mass of $10^{-5.50}$, consistent with the outward diffusion of helium over time.

@@

GD358 was discovered in 1982 \citep{Winget82} and has been the target of the Whole 
Earth Telescope (WET) in 1990, 1994, 2000, and 2006 \citep{Provencal09, Kepler03,Winget94}. 
New observations presented here include 278 individual observing runs (1105.1 hrs) spanning 
2007-2016 (Table \ref{journal}). Each season of observations was obtained as part of 
multi-site WET campaigns \citep{wet90}.

@@

\begin{figure}
\plotone{noise}
\caption{A comparison of the average noise as a function of frequency for the 2007-2016 observing 
seasons.  Each data set was prewhitened by its dominant frequencies.  The noise levels for each 
season are somewhat different.  This must be taken into account during frequency analysis.  
\label{noise}
}
\end{figure}

@@

Fig. \ref{schematic} also shows that the widths of the bands are not constant, but change 
from band to band. \citet{Bell15} presents an interesting analysis of the hydrogen atmosphere pulsator 
(DAV) KIC4552982, in which they identify 17 bands of pulsation frequencies. KIC4552982 is one of the 
coolest ZZ Cetis known \citep{Tremblay2013}, and it follows the general pattern exhibited 
by GD358: the highest frequency (shortest period) mode shows evidence of rotational splitting, while the 
lower frequency (longer period) modes are complex bands. The authors note that this DAV's bands have 
different widths in frequency space, and that there may be astronomical significance to this. 
Although the observational timebases and coverage are quite different (20 continuous months 
for KIC4552982 vs 34 incomplete years for GD358), our data presents the opportunity to investigate this 
for a cooler DB pulsator.  We measured the widths of each of GD358's band where we have
more than 10 detections. We define ''width'' as the difference between the lowest and highest frequency 
detected in each band. Fig. \ref{thick} presents the results. We find no correlation of width with number
of detected peaks in each band. Interestingly, the widths of GD358's bands are commensurate with those 
found in KIC4552982. We find a general increase in width with decreasing frequency (increasing period), 
until we reach the band at 1238 \muHz\ (807 s). This band has a width at least twice as wide as any other.

\citet{Mont2016} recently outlined a possible explanation for the behavior seen in the low frequency modes
of these cool DAVs that should also apply to GD358. White dwarfs pulsate in nonradial g-modes. As a DBV such as 
GD358 pulsates, it experiences local surface temperature variations as large as 3000 K.  The temperature 
variation affects different modes in different ways.  An appropriate analogy is to consider each mode's 
propagation region as a box with a lid. The box is defined as the region where the mode frequency is 
less than both the buoyancy (Brunt-V{\"a}is{\"a}l{\"a}) and acoustic (Lamb) frequencies.  The box's lid 
represents the mode's outer turning point. For short period modes (such as $k=8$ and $k=9$ in GD358), the 
lid is defined by the acoustic frequency, which is relatively insensitive to surface temperature variations. 
As the star pulsates, the box will not change, so these modes should be stable.  For the longer period modes, 
the lid of the box is defined by the buoyancy frequency, which goes to zero at the base of the surface 
convection zone. This is the important point: for longer period modes, the box lid
is actually defined by the base of the convection zone, which is very sensitive to local temperature variations.  
As GD358 pulsates, its convection zone deepens and thins in response to the local temperature variations. 
In our analogy, the lid of the box moves, effectively changing the characteristics of the box. Long period modes with 
outer turning points defined by the base of the convection zone should be perturbed and this is indeed 
what we see in GD358.

An additional interesting behavior is given by the distribution of average amplitudes for the bands 
(Fig~\ref{schematic}). In particular, the two bands at 1857.7 and 2007.6  \muHz\ have never been observed 
at amplitudes above 4 mma. The band at 1369 \muHz\ is also only observed at lower amplitudes. Interestingly, 
the band at 1741.5 \muHz\ was not observed at large amplitude prior to 2006, and it has remained at high 
amplitude since that time.

To summarize this section in broad brush strokes, the observed global pulsations involve the whole star, 
but each pulsation samples the star in slightly different ways.  Modes with lower frequencies 
(higher radial $k$ values) preferentially sample the outer layers, while modes with higher frequencies 
(lower radial $k$ values) have outer turning points that are farther from the surface, and so sample the 
deeper interior. It makes intuitive sense that lower frequency modes would be affected by processes 
confined to the outer stellar atmosphere, such as the convection zone. We speculate that the 
observed band widths and pulsation amplitudes contain information about the convection zone and/or any surface
magnetic field. Further investigation requires guidance from theory.

>>>>

\begin{abstract}
We report on the analysis of 33 years of photometric observations of the pulsating
helium atmosphere white dwarf GD358.  The complete data set includes archival data from 1982-2006, 
and 1105.1 hours of new observations from 2007-2015. From this data set, we extract 15 frequencies 
representing g-mode pulsation modes, adding 4 modes to the 11 modes 
known previously. We present evidence that these 15 modes are $\ell=1$ modes, 13 of which belong to 
a consecutive sequence in radial overtone
$k$. We perform a detailed asteroseismic analysis using models that include parameterized, complex carbon and oxygen 
core composition profiles to fit the periods. Recent spectroscopic analyses place GD358 near the red edge of the DBV instability strip, at 24,000 $\pm$ 500~K and a \logg of 7.8 $\pm$ 0.08 dex. The surface gravity translates to a mass range of 0.455 to 0.540 \msun. Our best fit model has a temperature of 23,650~K and a mass of 0.5706 \msun ~. That is slightly more massive than what the most recent spectroscopy suggests. We find a pure helium layer mass of $10^{-5.50}$, consistent with the outward diffusion of helium over time.

@@

GD358 was discovered in 1982 \citep{Winget82} and has been the target of the Whole 
Earth Telescope (WET) in 1990, 1994, 2000, and 2006 \citep{Provencal09, Kepler03,Winget94}. 
New observations presented here include 254 individual observing runs (1105.1 hrs) spanning 
2007-2015 (Table \ref{journal}). Each season of observations was obtained as part of 
multi-site WET campaigns \citep{wet90}.

@@

\begin{figure}
\plotone{noise}
\caption{A comparison of the average noise as a function of frequency for the 2007-2015 observing 
seasons.  Each data set was prewhitened by its dominant frequencies.  The noise levels for each 
season are somewhat different.  This must be taken into account during frequency analysis.  
\label{noise}
}
\end{figure}

@@

Fig. \ref{schematic} also shows that the widths of the bands are not constant, but change 
from band to band. \citet{Bell15} presents an interesting analysis of the hydrogen atmosphere pulsator 
(DAV) KIC4552982, in which they identify 17 bands of pulsation frequencies. KIC4552982 is one of the 
coolest ZZ Cetis known \citep{Tremblay2013}, and it follows the general pattern exhibited 
by GD358: the highest frequency (shortest period) mode shows evidence of rotational splitting, while the 
lower frequency (longer period) modes are complex bands. The authors note that this DAV's bands have 
different widths in frequency space, and that there may be astronomical significance to this. 
Although the observational timebases and coverage are quite different (20 continuous months 
for KIC4552982 vs 33 incomplete years for GD358), our data presents the opportunity to investigate this 
for a cooler DB pulsator.  We measured the widths of each of GD358's band where we have
more than 10 detections. We define ''width'' as the difference between the lowest and highest frequency 
detected in each band. Fig. \ref{thick} presents the results. We find no correlation of width with number
of detected peaks in each band. Interestingly, the widths of GD358's bands are commensurate with those 
found in KIC4552982. We find a general increase in width with decreasing frequency (increasing period), 
until we reach the band at 1238 \muHz\ (807 s). This band has a width at least twice as wide as any other.

A second interesting behavior is given by the distribution of average amplitudes for the bands 
(Fig~\ref{schematic}). In particular, the two bands at 1857.7 and 2007.6  \muHz\ have never been observed 
at amplitudes above 4 mma. The band at 1369 \muHz\ is also only observed at lower amplitudes. Interestingly, 
the band at 1741.5 \muHz\ was not observed at large amplitude prior to 2006, and it has remained at high 
amplitude since that time.

In broad brush strokes, the observed global pulsations involve the whole star, but each pulsation samples the star
in slightly different ways.  For example, modes with lower frequencies (higher radial $k$ values) preferentially 
sample the outer layers, while modes with higher frequencies (lower radial $k$ values) have outer turning points 
that are farther from the surface, and so sample the deeper interior. It makes intuitive sense that lower 
frequency modes would be affected by processes confined to the outer stellar atmosphere. We speculate that the 
observed band widths and pulsation amplitudes contain information about the convection zone and/or any surface
magnetic field. Further investigation requires guidance from theory.
